\textbf{Policy should change to improve invited speaker diversity and reflect trainee diversity}

### 2019-09-27 submission/manuscript.tex
revert legends
minor edits, prep for biorxiv submission

\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Policy should change to improve invited speaker diversity and retain trainees},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Policy should change to improve invited speaker diversity and
retain trainees}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

The biological sciences have a problem retaining white women and
underrepresented minorities in academia. Despite increases in the
representation of these groups in faculty candidate pools, they are
still underrepresented at the faculty level, particularly at the Full
professor level. Faculty are often invited to departments at other
universities to present on their research, a highly valued offer that
provides an opportunity for collaborations and networking. This is also
an opportunity for trainees to interact with additional faculty from
underrepresented groups. In this manuscript, we compare invited-speaker
demographics to the current trainee demographics and find that trainees
are not proportionally represented by speakers invited to the
department. Our investigation prompted changes in policy for how invited
speakers are selected in the future to invite a more diverse group of
scientists. To facilitate this process, we developed a set of tips and a
web-based resource that allows scientists, committees, and moderators to
identify members of under-served groups. These resources can be easily
adapted by other fields or sub-fields to promote inclusion and diversity
at seminar series', conferences, and colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the
under-representation of many racial and ethnic groups, as well as women,
in science, technology, engineering, and math (STEM) fields (NIH 2015
report, Measuring DEI 2016, (1)). Organization climate and culture that
supports inclusion of all individuals is essential to support the
retention of historically underrepresented minorities (HURM) in the
U.S., as well as non-Caucasian individuals, and women in STEM fields
(Schneider, 2013). Inclusive organization culture will not suffice in
retaining HURM, non-Caucasian individuals, and women without
organization policy changes that support their values and goals (2).

A long-standing issue within STEM fields, specifically within academia,
has been a lack of diverse representation of scientists for trainees
(graduate students and postdoctoral fellows). In order to maintain
retention of HURM, non-Caucasian individuals, and women in STEM fields,
it is important for trainees to have visual representations of
themselves as scientists. The importance of representation in retaining
a diverse group of individuals in STEM fields is supported by social
role theory (3). Individuals tend to make inferences about
characteristics that are needed to be successful in a given role by
examining individuals that most occupy that role (3, 4). Therefore,
trainees who do not see representation of themselves in senior
scientific positions, or in this case as senior faculty members, may
decide that they do not possess the characteristics that are required to
succeed. In an attempt to make science a more inclusive environment,
many individuals have attempted to address this issue by promoting the
inclusion of more women speakers at conferences (5--7). However, no
study to date has addressed invited speaker seminar series diversity
within their own institution.

Within the Department of Microbiology and Immunology at the University
of Michigan (Ann Arbor, MI, USA), each year from September to June,
faculty members have the opportunity to invite scientists from other
institutions to give a one-hour seminar in which all department members
attend. The invited seminar speaker also has the opportunity to meet
with current department faculty members and trainees. Scientists who are
invited to give seminars are widely regarded as successful and the top
in their field. Thus, if trainees are constantly being exposed to ``the
top scientist in their field'', according to social role theory, it is
imperative that the selected seminar speakers represent a diverse group
of individuals.

In this study, we examine and compare the proportion of HURM,
non-Caucasian/non-HURM, and women invited speakers to white males in the
Department of Microbiology and Immunology. Additionally, we compare
invited-speaker demographics to the current trainee demographics as a
means to gauge if trainee demographics are being represented accordingly
throughout the seminar series. Following our investigation, we proposed
a policy change to the Department of Microbiology and Immunology in how
invited speakers are selected as a means to promote inclusion in our
department and reduce stereotype threat, microaggressions, and
unconscious bias. In order to facilitate inviting a more diverse group
of scientists, we developed a set of resources that allow scientists,
within the field of microbiology, to self-identify as an HURM,
non-Caucasian/non-HURM, or a woman. These resources will promote
inclusion and diversity by providing greater representation of all
scientists and will provide hosts an opportunity to invite a more
diverse group of scientists.

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. The trainees were identified using
from departmental email lists that included masters students, doctoral
students, and post-doctoral fellows.

@@

Data were analyzed and figures generated in R Statistical Software,
using relevant packages (8--20).

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers (80\% each), relative to the
proportion of Caucasian trainees, which was 55\% (Fig. 1B). We also
observed declines in the representation of HURM and NCNH faculty and
speakers relative to the trainees (Fig 1B). HURM trainees made up 11\%
of the department, on track with the 11\% of microbiology and immunology
doctorates awarded in 2017 (21). However, only 8.5\% of invited
speakers, and none of the hosting faculty, were HURM scientists. NCNH
trainees were 34\% of department students and postdoc (22\% of
microbiology and immunology doctorates in 2017), but only 19\% of
hosting faculty and 10.5\% of invited speakers (21).

The more prestigious invited speaker lectureships were also dominated by
Caucasian scientists, who comprised 81.48\% of those awarded (Fig. 1C).
HURM and NCNH scientists were awarded 3 and 2 lectureships,
respectively. Because the intersection of identities can compound biases
and outcomes, we further examined the lectureships by gender and
race/ethnicity status. Caucasian men and women accounted for 44.44\% and
37.04\% of the lectureships, respectively. Just 18.52\% of lectureships
were held by non-Caucasian men while none were held by non-Caucasian
women (Fig. 1D).

@@

In our study population, the proportion of women as faculty hosts and
speakers is equivalent to global estimates that 40\% of microbiologists
are women, with a slightly lower representation of women in lectureships
(22). In that regard, women are over-represented as graduate students
and postdoctoral fellows in the department. Caucasian scientists are
over-represented as host faculty and invited speakers, compared to their
presence as trainees, particularly when lectureships were considered.
Conversely, the proportion of HURM and NCNH invited speakers were
under-representative of the trainee populations for each group. Thus,
according to social role theory, this department is not adequately
serving their non-Caucasian trainees.

Several papers have investigated the representation of women at
scientific conferences, however, we have only identified one that
focused on invited speakers at universities (23). In their study,
Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in 2013 -
2014 and found that women faculty are less likely to be invited
speakers, despite similar acceptance rates. These results suggest that
women faculty are less often invited as speakers. We have not been able
to identify any other publications examining scientific speaker
diversity beyond gender. This seems to be the first, which is concerning
since conclusions drawn from gender-based studies are often framed, and
considered, to be applicable to other marginalized groups (e.g., HURM).
This is a flawed assumption. While there is no doubt some overlap, each
group remains marginalized due to a unique complex set of factors that
cannot always be solved by gender-based solutions. US-serving
institutions, such as the University of Michigan have a particular
responsibility to the historically suppressed populations included in
our definition of HURMs. We therefore implore US institutions to apply
this framing to their discussions and research.

Departments have different processes and criteria for selecting invited
speakers, but it is a matter of pride to bring the best scientists
possible. It may be that the definition of ``best'' poses a problem to
women, who need three-times as many publications as their men colleges
to be considered equally competent (24), and those that identify as HURM
and Asian, who report having to work harder than white men to be
perceived as legitimate scholars (25). Some departments only invite
tenured faculty, which severely limits the number of potential speakers
who are white women or non-Caucasian. Yet, another scenario is that
pre-tenure faculty members invite prestigious, tenured faculty in their
field to network and secure letters for their own tenure package. The
increased burden of white women and non-Caucasian scientists to prove
competency decreases their likelihood to be considered for either tenure
or as possible source of tenure letters.

Each underrepresented group in our cohort faces a complex set of
barriers to achieving faculty status. For instance, the decision to
invite a woman may also be negatively impacted by assumptions about
competency and dedication. The dedication of women who have children to
their work is perceived to be less than that of their colleagues, i.e.,
men who also have children (26--28). The perceived prioritization and
commitments of women to family over work may cause faculty to doubt
their acceptance of a speaking invitation, despite the prestigious
nature of these invitations and evidence that men and women accept at
similar rates ({\textbf{???}}, 23). As a result, the faculty member
invites a different colleague who they feel is more likely to agree (and
is a man). Another large portion of our sample were the NCNH cohort, who
are predominately Asian/Asian American individuals. Although Asian
scientists are well-represented in the US scientific workforce, they
face significant bias and barriers to inclusion in society and academia
(29, 30). For instance, despite the higher employment rate of Asian
scientists, they were not well-represented in the more prestigious
lectureships.

While HURM and NCNH share some experiences, differences including
varying rates of hiring and tenure promotion mean unique considerations
are important for inclusion of each group (31). For instance, a major
barrier to inclusion of HURM faculty at similar proportions to HURM
trainees is the low transition rate of scientists from HURM backgrounds
to faculty positions and the associated low proportion of HURM faculty
(32). The proportion of HURM faculty at the Assistant and Associate
Professor level is currently higher than at Full Professor so it will be
difficult to increase speaker diversity if early-career researchers are
not being considered (33). Increased performance expectations and
patterns of exclusions are consistent themes in studies characterizing
the HURM faculty experience (34, 35). Therefore, inclusion of HURM
faculty in seminar series is likely essential to increasing the number
of HURM Associate and Full Professors. Even when HURM speaker rates
match the proportion of HURM faculty employment, HURM trainees will be
represented at a significantly higher proportion. Inclusion of HURM
faculty in these seminar series is just one aspect of larger
institutional change that is needed (36).

@@

In an attempt to promote inclusion within the Department of Microbiology
and Immunology at the University of Michigan, these data were presented
to faculty members and the department chair (Dr.~Mobley). Since trainee
demographics were not represented within seminar speaker demographics
over the past five years, we proposed a policy change as to how seminar
speakers were being invited. One suggestion was to switch from
faculty-invited to lab-invited speakers in an attempt to allow trainees
to choose a speaker that best represented themselves (Table 1).

The implicit biases that affect perceptions of marginalized groups are
an issue, but we must acknowledge that it is not always possible to
identify members of historically under-served communities. For instance,
after data analysis, we learned that at least one speaker in our data
set should have been categorized as a HURM instead of Caucasian, but it
wasn't readily apparent from their internet presence or CV. This
limitation makes two important points: that perceived identity often
plays a larger role than self-identification, and that we need better
tools to identify members of marginalized groups. Another policy
suggestion is for departments to invite their speakers to spend a time
discussing their personal journeys through science, in addition to their
scientific stories (Table 1). This would enable those who wish, feel
welcomed to discuss how their identity(ies) interacted with their
careers. In addition to these suggestions for policy change, we have
created resources that allow scientists to self-identify as under-served
groups and thus provide host faculty with more diverse choices (Table
1).

@@

Motivated by a lack of resources to identify scientists who are members
of marginalized and/or historically under-served groups, and inspired by
resources in other fields--DiversifyEEB and DiversifyChemistry--we
created DiversifyMicrobiology and DiversifyImmunology. These resources
are a tool for symposium organizers, award committees, search
committees, and other scientists to identify individuals to diversify
their pools. Additionally, we have built these as a template to be used
as a tool for other fields and organizations that wish to create their
own lists. Since these lists are compiled by self-nomination, we can
ensure that only scientists comfortable revealing their marginalized
identities are included.

@@

The website provides an interface to the Google forms and spreadsheets
with template pages for viewing the list, adding a name to the list, and
finding additional resources. Importantly, our website creation tool is
hosted for free by GitHub, which provides a free website for each GitHub
organization. Basic tools and skills required to set up a Diversify site
include knowledge of, or experience with, the version control tool git,
the web-tool GitHub, and a text editor. A tutorial in the
DiversifyMicrobiology repository on GitHub provides links to these
resources and instructions for adapting the tool to your own field.

@@

A.K.H. collected the data, assigned demographics, analyzed the data, and
created the website. R.M.P. created the Google lists, forms, and website
content and the description of their maintenance. J.L. wrote the
introduction and provided conceptual advice. A.K.H. and J.L facilitated
the policy change to the UM Department of Microbiology and Immunology.
All authors contributed to preparing the final manuscript.

@@

The anonymized data, code for all analysis steps, and an Rmarkdown
version of this manuscript is available at
\url{https://github.com/akhagan/Hagan_SpeakerDiversity_XXXX_2019/}.
Template and complete instructions for generating a field-specific
Diversity website are available at
\url{https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/}.

@@

\hypertarget{refs}{}
\hypertarget{ref-colloquy_2012}{}
1. 2012. Colloquy on minority males in science, technology, engineering,
and mathematics. National Academies Press.

\hypertarget{ref-Coe2019}{}
2. \textbf{Coe IR}, \textbf{Wiley R}, \textbf{Bekker L-G}. 2019.
Organisational best practices towards gender equality in science and
medicine. The Lancet \textbf{393}:587--593.
doi:\href{https://doi.org/10.1016/s0140-6736(18)33188-x}{10.1016/s0140-6736(18)33188-x}.

\hypertarget{ref-Eagly1984}{}
3. \textbf{Eagly AH}, \textbf{Steffen VJ}. 1984. Gender stereotypes stem
from the distribution of women and men into social roles. Journal of
Personality and Social Psychology \textbf{46}:735--754.
doi:\href{https://doi.org/10.1037/0022-3514.46.4.735}{10.1037/0022-3514.46.4.735}.

\hypertarget{ref-Carter2018}{}
4. \textbf{Carter AJ}, \textbf{Croft A}, \textbf{Lukas D},
\textbf{Sandstrom GM}. 2018. Women's visibility in academic seminars:
Women ask fewer questions than men. PLOS ONE \textbf{13}:e0202743.
doi:\href{https://doi.org/10.1371/journal.pone.0202743}{10.1371/journal.pone.0202743}.

\hypertarget{ref-kalejta_gender_2017}{}
5. \textbf{Kalejta RF}, \textbf{Palmenberg AC}. 2017. Gender Parity
Trends for Invited Speakers at Four Prominent Virology Conference
Series. Journal of Virology \textbf{91}.
doi:\href{https://doi.org/10.1128/JVI.00739-17}{10.1128/JVI.00739-17}.

\hypertarget{ref-casadevall_presence_2014}{}
6. \textbf{Casadevall A}, \textbf{Handelsman J}. 2014. The Presence of
Female Conveners Correlates with a Higher Proportion of Female Speakers
at Scientific Symposia. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mBio.00846-13}{10.1128/mBio.00846-13}.

>>>>

\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Policy should change to improve invited speaker diversity and reflect trainee diversity},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Policy should change to improve invited speaker diversity and
reflect trainee diversity}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

The biomedical sciences have a problem retaining white women and
underrepresented minorities in academia. Despite increases in the
representation of these groups in faculty candidate pools, they are
still underrepresented at the faculty level, particularly at the Full
Professor level. The lack of diverse individuals at the Full Professor
level contributes to the attrition of women and under-represented
minorities, as it confirms unconscious biases. The presence of
unconscious biases contribute to feelings of not belonging by trainees
and are amplified by visual representation of who is presented as the
``top scientist in their field''. Top scientists are not only defined by
the attainment of Full Professorships, but also through invited seminar
series. Invitations for faculty to present their research at other
university departments is highly valued offer that provides an
opportunity for collaborations and networking. However, if invited
speakers do not represent the demographics of current trainees, these
visual representations of successful scientists may contribute to
decreased attitudes of self-identification as a scientist, ultimately
resulting in trainees leaving the field or the academy. In this study,
we compare invited-speaker demographics to the current trainee
demographics in one microbiology and immunology department and find that
trainees are not proportionally represented by speakers invited to the
department. Our investigation prompted changes in policy for how invited
speakers are selected in the future to invite a more diverse group of
scientists. To facilitate this process, we developed a set of tips and a
web-based resource that allows scientists, committees, and moderators to
identify members of under-served groups. These resources can be easily
adapted by other fields or sub-fields to promote inclusion and diversity
at seminar series', conferences, and colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the
under-representation of many racial and ethnic groups, as well as women,
in science, technology, engineering, and math (STEM) fields (1--4).
Specifically, within the field of biomedical research in the United
States, the proportion of underrepresented minorities at the full
professor level has remained consistently low at 4\% (survey data taken
from the NIH from 2001 to 2013), compared to the U.S. population, which
is 32.3\% (5, 6). Similar discrepancies exist for women in biomedical
sciences as full professorships are currently held mostly by men (7, 8).
As demographics of faculty within the biomedical sciences remains skewed
towards Caucasian men, the demographics of trainees (graduate students
and postdocs) are becoming more diverse (5).

Policy changes are needed to support inclusion of all individuals,
particularly in the biomedical sciences(9). To increase retention of
historically under-represented minorities (HURM), non-Caucasian/non-HURM
(NCNH) individuals, and white women in biomedical fields, it is
important for trainees to have visual representations of themselves as
scientists. The importance of representation in retaining a diverse
group of individuals in STEM fields is supported by social role theory
(10). Individuals make inferences about characteristics that are needed
to be successful in a given role by examining individuals that most
occupy that role (10, 11). However, there is a lack of diverse
scientific experts in academia so underrepresented minorities are not
seeing adequate visual representations of themselves at the faculty
level. Therefore, trainees who do not see representation of themselves
in senior faculty positions, may decide that they do not possess the
characteristics that are required to succeed.

Invited seminar series are common within biomedical departments across
the United States (12). Usually, seminar series' consist of faculty
members selecting a scientist from another institution to visit their
university and present their research, as well as meet with other
faculty members and trainees. Named lectureships follow the same format
but are decided by committee and are considered more prestigious because
they are named in honor of prominent local scientists. These seminar
series and lectureships provide an opportunity for trainees to be
exposed to research outside of their department. Additionally, being an
invited speaker provides the scientist with an opportunity to make
future collaborations and build their own \emph{curriculum vitae} (CV).
Scientists invited to give seminars are widely regarded as successful
and the top in their field. Thus, if trainees are constantly being
exposed to ``the top scientist in their field'', according to social
role theory, it signals who is successful in that field. While some have
examined this issue by studying and promoting the inclusion of more
women speakers at conferences, how department speaker series compare to
the trainee diversity of that department is unknown (13--15).

In this study, we examine and compare the proportion of HURM, NCNH, and
women invited speakers to white men in the Department of Microbiology
and Immunology at the University of Michigan. Additionally, we compare
invited-speaker demographics to the current trainee demographics as a
means to gauge if trainee demographics are being represented accordingly
throughout the seminar series. Following our investigation, we proposed
a policy change to the Department of Microbiology and Immunology in how
invited speakers are selected as a means to promote inclusion in our
department and reduce unconscious bias. In order to facilitate inviting
a more diverse group of scientists, we developed a set of resources that
allow scientists, within the fields of microbiology and immunology, to
self-identify as having an under-represented or under-served identity
including: HURM, non-Caucasian/non-HURM, or a white woman. These
resources will promote inclusion and diversity by providing greater
representation of all scientists and will provide hosts an opportunity
to invite a more diverse group of scientists.

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. The trainees were identified using
departmental email lists that included masters students, doctoral
students, and post-doctoral fellows.

@@

Data were analyzed and figures generated in R Statistical Software,
using relevant packages (16--28).

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers (80\% each), relative to the
proportion of Caucasian trainees, which was 55\% (Fig. 1B). We also
observed declines in the representation of HURM and NCNH faculty and
speakers relative to the trainees (Fig 1B). HURM trainees made up 11\%
of the department, on track with the 11\% of U.S. microbiology and
immunology doctorates awarded in 2017 (29). However, only 8.5\% of
invited speakers, and none of the hosting faculty, were HURM scientists.
NCNH trainees were 34\% of department students and postdocs (versus 22\%
of U.S. microbiology and immunology doctorates in 2017), but only 19\%
of hosting faculty and 10.5\% of invited speakers (29).

The more prestigious invited speaker lectureships were also dominated by
Caucasian scientists, who comprised 81.48\% of those awarded (Fig. 1C).
HURM and NCNH scientists were awarded 3 and 2 lectureships,
respectively. Because the intersection of identities can compound biases
and outcomes, we further examined the lectureships by gender and
race/ethnicity status (30). Caucasian men and women accounted for
44.44\% and 37.04\% of the lectureships, respectively. Just 18.52\% of
lectureships were held by non-Caucasian men while none were held by
non-Caucasian women (Fig. 1D).

@@

This study found that the proportion of HURM and NCNH invited speakers
were under-representative of the trainee populations for each group.
Additionally, within the last 5 years, no HURM or NCNH woman was awarded
a lectureship. This means that the department is not providing
non-Caucasian trainees with adequate representation of successful
scientists. Taking this into context of social role theory, by not
adequately representing the diversity of all trainees, the department is
not supporting an inclusive environment in terms of visual faculty
representation. We also found that the proportion of women as faculty
hosts and speakers in our study population is equivalent to global
estimates that 40\% of microbiologists are women, though women only
represent about 30\% of academic biomedical faculty (7, 31). Women are
also over-represented as graduate students and postdoctoral fellows in
this department. Overall, Caucasian scientists are over-represented as
host faculty and invited speakers, compared to their presence as
trainees, particularly when lectureships were considered.

Several papers have investigated the representation of women at
scientific conferences, however, we have only identified one that
focused on invited speakers at universities (12). In their study,
Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in 2013 -
2014 and found that women faculty are less likely to be invited
speakers, despite similar acceptance rates (12). We have not been able
to identify any publications examining scientific speaker diversity
beyond gender. This seems to be the first, which is concerning since
conclusions drawn from gender-based studies are often framed, and
considered, to be applicable to other marginalized groups (e.g., HURM).
This is a flawed assumption. While there is no doubt some overlap, each
group remains marginalized due to a unique complex set of factors that
cannot always be solved by gender-based solutions. U.S. institutions,
such as the University of Michigan have a particular responsibility to
the historically suppressed populations included in our definition of
HURMs. We therefore implore U.S. institutions to apply this framing to
their discussions and research.

Departments have different processes and criteria for selecting invited
speakers, but it is a matter of pride to bring the best scientists
possible. It may be that the definition of ``best'' poses a problem to
under-represented and under-served groups (e.g., white women, HURM, and
Asian) who are held to stricter competency standards and report having
to work harder than white men to be perceived as legitimate scholars
(32, 33). Some departments only invite tenured faculty, which severely
limits the number of potential speakers who are white women or
non-Caucasian. Yet, another scenario is that pre-tenure faculty members
invite prestigious, tenured faculty in their field to network and secure
letters for their own tenure package. The increased burden of white
women and non-Caucasian scientists to prove competency decreases their
likelihood to be considered for either tenure or as possible source of
tenure letters.

Each underrepresented group in our cohort faces a complex set of
barriers to achieving faculty status. For instance, the decision to
invite a woman may also be negatively impacted by assumptions about
competency and dedication. The dedication of women who have children to
their work is perceived to be less than that of their colleagues, i.e.,
men who also have children (34--36). The perceived prioritization and
commitments of women to family over work may cause faculty to doubt
their acceptance of a speaking invitation, despite the prestigious
nature of these invitations and evidence that men and women accept at
similar rates (12, 37). As a result, the faculty member may invite a
different colleague who they feel is more likely to agree (and is a
man). Another large portion of our sample were the NCNH cohort, who are
predominately Asian/Asian American individuals. Although Asian
scientists are well-represented in the US scientific workforce, they
face significant bias and barriers to inclusion in society and academia
(38, 39). For instance, despite the higher employment rate of Asian
scientists, they were not well-represented in the more prestigious
lectureships.

While HURM and NCNH share some experiences, differences including
varying rates of hiring and tenure promotion mean unique considerations
are important for inclusion of each group (3). For instance, a major
barrier to inclusion of HURM faculty at similar proportions to HURM
trainees is the low transition rate of scientists from HURM backgrounds
to faculty positions and the associated low proportion of HURM faculty
(40). The proportion of HURM faculty at the Assistant and Associate
Professor level is currently higher than at Full Professor so it will be
difficult to increase speaker diversity if early-career researchers are
not being considered (41). Increased performance expectations and
patterns of exclusions are consistent themes in studies characterizing
the HURM faculty experience (42, 43). Therefore, inclusion of HURM
faculty in seminar series is likely essential to increasing the number
of HURM Associate and Full Professors. Even when HURM speaker rates
match the proportion of HURM faculty employment, HURM trainees will be
represented at a significantly higher proportion. Inclusion of HURM
faculty in these seminar series is just one aspect of larger
institutional change that is needed (44).

@@

In an attempt to promote inclusion within the Department of Microbiology
and Immunology at the University of Michigan, these data were presented
to faculty members and the department chair (Dr.~Mobley). Since trainee
demographics were not represented by the seminar speaker demographics
over the past 5 years, we proposed a policy change as to how seminar
speakers were being invited. One suggestion was to switch from
faculty-invited to lab-invited speakers in an attempt to allow trainees
to choose a speaker that best represented themselves (Table 1).

The implicit biases that affect perceptions of marginalized groups are
an issue, but we must acknowledge that it is not always possible to
identify members of historically under-served communities. For instance,
after data analysis, we learned that at least one speaker in our data
set should have been categorized as a HURM instead of Caucasian, but it
wasn't readily apparent from their internet presence or CV. This
limitation makes two important points: that perceived identity often
plays a larger role than self-identification, and that we need better
tools to identify members of marginalized groups. Another policy
suggestion is for departments to invite their speakers to spend time
discussing their personal journeys through science, in addition to their
scientific stories (Table 1). This would enable those who wish, to
discuss how their identity(ies) interacted with their careers. In
addition to these suggestions for policy change, we have created
resources that allow scientists to self-identify as under-served groups
and thus provide host faculty with more diverse choices (Table 1).

@@

Motivated by a lack of resources to identify scientists who are members
of marginalized and/or historically under-served groups, and inspired by
resources in other fields--DiversifyEEB and DiversifyChemistry--we
created DiversifyMicrobiology and DiversifyImmunology (45--48). These
resources are a tool for symposium organizers, award committees, search
committees, and other scientists to identify individuals to diversify
their pools. Additionally, we have built these as a template to be used
by other fields and organizations that wish to create their own lists.
Since these lists are compiled by self-nomination, we can ensure that
only scientists comfortable revealing their marginalized identities are
included.

@@

The website provides an interface to the Google forms and spreadsheets
with template pages for viewing the list, adding a name to the list, and
finding additional resources. Importantly, our website creation tool is
hosted for free by GitHub, which provides a free website for each GitHub
organization. Basic tools and skills required to set up a Diversify site
include knowledge of, or experience with, the version control tool git,
the web-tool GitHub, and a text editor. A tutorial in the
DiversifyMicrobiology repository on GitHub provides links to these
resources and instructions for adapting the tool to your own field (47).

@@

A.K.H. collected the data, assigned demographics, analyzed the data,
created the website, and wrote the methods and results. R.M.P. created
the Google lists, forms, and website content and the description of
their maintenance. J.L. wrote the introduction and provided conceptual
advice. A.K.H. and J.L facilitated the policy change to the UM
Department of Microbiology and Immunology. All authors contributed to
preparing the final manuscript.

@@

The anonymized data, code for all analysis steps, and an Rmarkdown
version of this manuscript is available at
\url{https://github.com/akhagan/Hagan_SpeakerDiversity_JMBE_2019}.
Template and complete instructions for generating a field-specific
Diversity website are available at
\url{https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/}.

@@

\hypertarget{refs}{}
\hypertarget{ref-Martinez2018}{}
1. \textbf{Martinez LR}, \textbf{Boucaud DW}, \textbf{Casadevall A},
\textbf{August A}. 2018. Factors contributing to the success of
NIH-designated underrepresented minorities in academic and nonacademic
research positions. CBELife Sciences Education \textbf{17}:ar32.
doi:\href{https://doi.org/10.1187/cbe.16-09-0287}{10.1187/cbe.16-09-0287}.

\hypertarget{ref-AllenRamdial2014}{}
2. \textbf{Allen-Ramdial S-AA}, \textbf{Campbell AG}. 2014. Reimagining
the pipeline: Advancing STEM diversity, persistence, and success.
BioScience \textbf{64}:612--618.
doi:\href{https://doi.org/10.1093/biosci/biu076}{10.1093/biosci/biu076}.

\hypertarget{ref-Fang2000}{}
3. \textbf{Fang D}. 2000. Racial and ethnic disparities in faculty
promotion in academic medicine. JAMA \textbf{284}:1085.
doi:\href{https://doi.org/10.1001/jama.284.9.1085}{10.1001/jama.284.9.1085}.

\hypertarget{ref-Gibbs2014}{}
4. \textbf{Gibbs KD}, \textbf{McGready J}, \textbf{Bennett JC},
\textbf{Griffin K}. 2014. Biomedical science ph.D. career interest
patterns by race/ethnicity and gender. PLoS ONE \textbf{9}:e114736.
doi:\href{https://doi.org/10.1371/journal.pone.0114736}{10.1371/journal.pone.0114736}.

\hypertarget{ref-Meyers2018}{}
5. \textbf{Meyers LC}, \textbf{Brown AM}, \textbf{Moneta-Koehler L},
\textbf{Chalkley R}. 2018. Survey of checkpoints along the pathway to
diverse biomedical research faculty. PLOS ONE \textbf{13}:e0190606.
doi:\href{https://doi.org/10.1371/journal.pone.0190606}{10.1371/journal.pone.0190606}.
